Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-05-19 paper.tex
Edit down some lines.

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a sequence of primitive shape-drawing commands that can generate an image similar to the observed image. We refer to this sequence as a \emph{execution trace}, since it corresponds to the sequence of primitive commands that a program would have issued, but it lacks the high-level structure that determines how the program decided to issue those commands. For added robustness, we train the network to produce a proposal distribution of the most likely traces, and use stochastic search to find the trace that best matches the input. The network is trained from an automatically generated corpus of synthetic image generating programs.

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Figure~\ref{handDrawingExamples}; see supplement for details).
Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{equation}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{equation}
Intuitively, Eq.~\ref{symmetricDistance} says that
$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
the trace we want and the trace we have so far.  Pixel-wise distance
metrics are sensitive to the fine details of how and exactly where
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details. So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn reasonably carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid,
we made the drawings on graph paper.
\begin{figure}\centering
  \begin{minipage}{0.45\textwidth}
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}%
   \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
  \end{minipage}\hfill
  \begin{minipage}{0.45\textwidth}
  \includegraphics[width = 6cm]{figures/drawingAccuracy.png}
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1,5,10,100)?
  Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces (sets of drawing commands).}\label{drawingIntersectionOverUnion}
    \end{minipage}
\end{figure}

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-0.png}
    \includegraphics[width = 2cm]{figures/0-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}    
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}. Stochastic search (SMC) can help correct for these errors, as can the program synthesizer (Section~\ref{synthesizerHelpsParsing})}\label{lotsOfHandDrawings}%~\remark{Showing `failure cases' (the last three above) out-of-context seems not great. Perhaps remind people in the caption that stochastic search can help correct some of these issues?}
\end{figure}
For each drawing we annotated a ground truth trace, and evaluated the
model by asking it to sample many candidate traces for each drawing.
For 58\% of the drawings the Top-1 most likely sample exactly matches the
ground truth; as we consider more samples the model encounters traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because our current model sometimes makes mistakes on hand drawings,
we envision the current system working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

\section{Synthesizing graphics programs from execution traces}\label{programSynthesisSection}
Although the execution trace of a graphics program describes the parts
of a scene, it fails to encode higher-level features of the image,
such as repeated motifs or symmetries.  A \emph{graphics
  program} better describe structures like these,
and we now take as our goal to synthesize simple graphics programs from
their execution traces.

@@

The constrained optimization problem in
equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to
program synthesis problems like these. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see~\cite{solar2008program}.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem, and then uses a SAT solver to find a
minimum cost program satisfying the specification.  In exchange for
not having any guarantees on how long it will take to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.

\newcommand{\exampleProgramSize}{2.3cm} 
\begin{figure}
  \begin{tabular}{lllll}
  \includegraphics[width = \exampleProgramSize]{figures/expert-29.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-52.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-38.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-72.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-75.png}\\
    \begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize = \small]
for(i<3){
line(i,-1*i+6,
2*i+2,-1*i+6)
line(i,-2*i+4,
i,-1*i+6)
\end{Verbatim}
    \end{minipage}&
    \begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize = \small]
circle(4,10)
for(i<3){
circle(-3*i+7,5)
circle(-3*i+7,1)
line(-3*i+7,4,
-3*i+7,2,
arrow=True)
line(4,9,
-3*i+7,6,
arrow=True)
\end{Verbatim}
\end{minipage}&
\begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize=\small]
for(i<3)
for(j<3)
if(j>0)
line(-3*j+8,
-3*i+7,-3*j+9,
-3*i+7)
line(-3*i+7,-3*j+8,
-3*i+7,
-3*j+9)
circle(-3*j+7,-3*i+7)
\end{Verbatim}
\end{minipage}&
\begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize=\small]
reflect(y=8){
for(i<3){
if(i>0){
rectangle(3*i-1,
2,3*i,3)}
circle(3*i+1,
       3*i+1)
\end{Verbatim}
\end{minipage}&
 \begin{minipage}[t]{\exampleProgramSize}
    \begin{Verbatim}[fontsize = \small]
for(i<4){
line(-4*i+13,4,
-4*i+13,2,
arrow=True)
for(j<3){
if(j>0){
circle(-4*i+13,
4*j+-3)}
line(-4*j+10,5,
-4*j+12,5,
arrow=True)
\end{Verbatim}
  \end{minipage}
  \end{tabular}
  \caption{Example drawings (top row) and the programs we synthesize from their ground truth traces (bottom row). Notice the nested loops in the Ising model (middle), special case conditionals for the HMM (rightmost), combination of symmetry and iteration in middle left, and affine transformation in the leftmost figure.}
  \end{figure}

@@

\subsection{Correcting errors made by the neural network}\label{synthesizerHelpsParsing}
The program synthesizer can help correct errors from the execution
trace proposal network by favoring execution traces which lead to more
concise or general programs.  For example, one generally prefers
figures with perfectly aligned objects over figures whose parts are
slightly misaligned -- and precise alignment lends itself to short
programs.  Concretely, we estimated a prior over programs. Then,
given the few most likely traces output by the neurally guided sampler,
we reranked them according to the prior probability of their programs.
Our neurally guided sampler could only do better on
7 drawings by looking at the Top-100 sampled traces
(see Fig.~\ref{drawingIntersectionOverUnion}),
precluding a statistically significant analysis of how much
learning a prior over programs could help correct errors.
But,
learning this prior does sometimes
help correct mistakes made by the neural network, and also
occasionally introduces mistakes of its own; see
Fig.~\ref{exampleOfProgramCorrectingMistake} for a representative
example of the kinds of corrections that it makes.
See supplement for details.
\begin{wrapfigure}{R}{5.5cm}
  \includegraphics[width = 5cm]{figures/programSuccess7.png}
  \caption{Left: hand drawing. Center: interpretation favored by the deep network. Right: interpretation favored after learning a prior over programs. Our learned prior favors shorter, simpler programs, thus continuing the pattern of not having an arrow is preferred.}\label{exampleOfProgramCorrectingMistake}
  \end{wrapfigure}

@@

\subsection{Modeling similarity between drawings}
Modeling an image using a program opens up new ways of measuring similarity between drawings.
For example, we might say that two drawings are similar if they both contain repetitions of length 4,
or if they share the same reflectional symmetry,
or if they are both organized according to a grid-like structure.

@@

In the not-too-distant future, we believe it should be possible to produce professional-looking figures just by drawing them and then letting an artificially-intelligent agent write the corresponding code.
More generally, we believe that the two-phase system we have proposed---parsing into execution traces, then searching for a low-cost symbolic program which generates those traces---may be a useful paradigm for other domains in which agents must programmatically reason about noisy perceptual input.

>>>>

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a sequence of primitive shape-drawing commands that generates an image similar to the observed image. We refer to this sequence as a \emph{execution trace}, since it corresponds to the sequence of primitive commands that a program would have issued but lacks the high-level structure that determines how the program decided to issue them. For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the input. The network is trained from an automatically-generated corpus of synthetic image-generating programs.

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Figure~\ref{handDrawingExamples}; see supplement for details).
Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{equation}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{equation}
Intuitively,
$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
the trace we want and the trace we have so far.  Pixel-wise distance
metrics are sensitive to the fine details of how and exactly where
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details. So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn reasonably carefully but not perfectly.
Because our model assumes that objects lie on a $16\times 16$ grid,
we made the drawings on graph paper.
\begin{figure}\centering
  \begin{minipage}{0.45\textwidth}
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}%
   \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.3\textwidth}\includegraphics[width = 1.5cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
  \end{minipage}\hfill
  \begin{minipage}{0.45\textwidth}
  \includegraphics[width = 6cm]{figures/drawingAccuracy.png}
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1,5,10,100)?
  Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces (sets of drawing commands).}\label{drawingIntersectionOverUnion}
    \end{minipage}
\end{figure}

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-0.png}
    \includegraphics[width = 2cm]{figures/0-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}    
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}. Stochastic search (SMC) can help correct for these errors, as can the program synthesizer (Section~\ref{synthesizerHelpsParsing})}\label{lotsOfHandDrawings}%~\remark{Showing `failure cases' (the last three above) out-of-context seems not great. Perhaps remind people in the caption that stochastic search can help correct some of these issues?}
\end{figure}
For each drawing we annotated a ground truth trace and asked the model
to sample many candidate traces for each drawing.
For 58\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
As the model sometimes makes mistakes on hand drawings,
we envision the current system working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

\section{Synthesizing graphics programs from execution traces}\label{programSynthesisSection}
While the execution trace of a graphics program describes the contents
of a scene, it does not encode higher-level features of the image,
such as repeated motifs or symmetries.  A \emph{graphics
  program} better describe such structures;
we seek to synthesize simple graphics graphics from their execution traces.

@@

The constrained optimization problem in
Equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see~\cite{solar2008program}.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.

\newcommand{\exampleProgramSize}{2.3cm} 
\begin{figure}
  \begin{tabular}{lllll}
  \includegraphics[width = \exampleProgramSize]{figures/expert-29.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-52.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-38.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-72.png}&
  \includegraphics[width = \exampleProgramSize]{figures/expert-75.png}\\
    \begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize = \small]
for(i<3){
line(i,-1*i+6,
2*i+2,-1*i+6)
line(i,-2*i+4,
i,-1*i+6)
\end{Verbatim}
    \end{minipage}&
    \begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize = \small]
circle(4,10)
for(i<3){
circle(-3*i+7,5)
circle(-3*i+7,1)
line(-3*i+7,4,
-3*i+7,2,
arrow=True)
line(4,9,
-3*i+7,6,
arrow=True)
\end{Verbatim}
\end{minipage}&
\begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize=\small]
for(i<3)
for(j<3)
if(j>0)
line(-3*j+8,
-3*i+7,-3*j+9,
-3*i+7)
line(-3*i+7,-3*j+8,
-3*i+7,
-3*j+9)
circle(-3*j+7,-3*i+7)
\end{Verbatim}
\end{minipage}&
\begin{minipage}[t]{\exampleProgramSize}
\begin{Verbatim}[fontsize=\small]
reflect(y=8){
for(i<3){
if(i>0){
rectangle(3*i-1,
2,3*i,3)}
circle(3*i+1,
       3*i+1)
\end{Verbatim}
\end{minipage}&
 \begin{minipage}[t]{\exampleProgramSize}
    \begin{Verbatim}[fontsize = \small]
for(i<4){
line(-4*i+13,4,
-4*i+13,2,
arrow=True)
for(j<3){
if(j>0){
circle(-4*i+13,
4*j+-3)}
line(-4*j+10,5,
-4*j+12,5,
arrow=True)
\end{Verbatim}
  \end{minipage}
  \end{tabular}
  \caption{Example drawings (top) and programs synthesized from their ground truth traces (bottom). Note the nested loops in the Ising model (middle), special case conditionals for the HMM (rightmost), combination of symmetry and iteration in middle left, and affine transformation in the leftmost figure.}
  \end{figure}

@@

\subsection{Correcting errors made by the neural network}\label{synthesizerHelpsParsing}
The program synthesizer can correct errors from the execution
trace proposal network by favoring traces which lead to more
concise or general programs.  For example, figures with perfectly aligned objects are preferrale to figures whose parts are slightly misaligned, and precise alignment lends itself to short
programs.  Concretely, we estimated a prior over programs. Then,
given the few most likely traces output by the neurally guided sampler,
we reranked them according to the prior probability of their programs.
Our sampler could only do better on
7 drawings by looking at the Top-100 sampled traces
(see Fig.~\ref{drawingIntersectionOverUnion}),
precluding a statistically significant analysis of how much
learning a prior over programs could help correct errors.
But,
learning this prior does sometimes
help correct mistakes made by the neural network, and also
occasionally introduces mistakes of its own; see
Fig.~\ref{exampleOfProgramCorrectingMistake} for a representative
example of the kinds of corrections that it makes.
See supplement for details.
\begin{wrapfigure}{R}{5.5cm}
  \includegraphics[width = 5cm]{figures/programSuccess7.png}
  \caption{Left: hand drawing. Center: interpretation favored by the deep network. Right: interpretation favored after learning a prior over programs. Our learned prior favors shorter, simpler programs, thus continuing the pattern of not having an arrow is preferred.}\label{exampleOfProgramCorrectingMistake}
  \end{wrapfigure}

@@

\subsection{Modeling similarity between drawings}
Modeling drawings using programs opens up new ways to measure similarity them.
For example, we might say that two drawings are similar if they both contain loops of length 4,
or if they share a reflectional symmetry,
or if they are both organized according to a grid-like structure.

@@

In the near future, we believe it will be possible to produce professional-looking figures by drawing them and then letting an artificially-intelligent agent write the corresponding code.
More generally, we believe that the two-phase system we have proposed---parsing into execution traces, then searching for a low-cost symbolic program which generates those traces---may be a useful paradigm for other domains in which agents must programmatically reason about noisy perceptual input.
